---
title: "MP3 - Sequels w/ SQL"
author: Audrey Bertin & Eva Gerstle
output: html_document
---

```{r, message = FALSE, warning = FALSE}

library(mdsr)
library(tidyverse)
library(RMySQL)
db <-dbConnect_scidb(dbname = "imdb")

```

In this project, we were looking to study movie sequels, remakes, and reboots throughout history. Our goal was to determine: 
  1) Whether the number of years between sequels and their originals is changing, and 
  if so, whether they're getting further apart or closer together.
  2) Whether there are more of these movies now than in the past


### SQL Queries to find the number of years between a movie and it's remake/sequel

```{r}

remakes <- db %>% 
  dbGetQuery(
    "SELECT ml.movie_id, 
    ml.linked_movie_id, 
    link_type_id, 
    t2.title AS remake, 
    t.title AS original, 
    t2.production_year AS remake_year,
    t.production_year AS original_year,
    (t2.production_year - t.production_year ) AS years_between
    FROM movie_link AS ml
    JOIN title AS t ON t.id = ml.movie_id
    JOIN title AS t2 ON t2.id = ml.linked_movie_id
    WHERE ml.link_type_id =  4
    AND t.kind_id = 1
    AND t2.kind_id = 1")
sequels <- db %>%
  dbGetQuery(
    "SELECT t2.title AS sequel, 
    t.title AS original, 
    t2.production_year AS sequel_year,
    t.production_year AS original_year,
    (t2.production_year - t.production_year ) AS years_between
    FROM movie_link AS ml
    JOIN title AS t ON t.id = ml.movie_id
    JOIN title AS t2 ON t2.id = ml.linked_movie_id
    WHERE ml.link_type_id =  2
    AND t.kind_id = 1
    AND t2.kind_id = 1")

#These queries both join twice on the primary key which is a very quick opperation. Due to the selecting of specific columns, SQL is further optimized. The remake query is much faster due to there being less remakes than sequels.  
```
### Finding the average number of years between an original and it's sequel/remake for each year
```{r}

avg_length <- remakes %>%
  group_by(remake_year) %>%
  summarize(avg_gap = mean(years_between))

avg_length_sequels <- sequels %>%
  group_by(sequel_year) %>%
  summarize(avg_gap = mean(years_between))

```
### Graphing the average gap between sequels/remakes  and originals for each year 

```{r, warning = FALSE}

avg_gap_sequels <- ggplot(avg_length_sequels, aes(x=sequel_year, y= avg_gap)) + geom_area(fill = "#e41a1c", alpha = 0.6) + scale_x_continuous(name='Year of Sequel', breaks = c(1900, 1910, 1920,1930, 1940, 1950, 1960,1970,1980,1990,2000,2010,2020), labels = c("1900", "1910","1920","1930","1940","1950","1960","1970","1980","1990","2000","2010","2020")) + scale_y_continuous(name ='Average years between Original and Sequel', limits = c(0,60)) + ggtitle("Spacing between Movies and their Sequels") 
avg_gap_sequels

avg_gap_remakes <- ggplot(avg_length, aes(x=remake_year, y= avg_gap)) + geom_area(fill = "#377eb8", alpha = 0.6) + scale_x_continuous(name='Year of Remake', breaks = c(1900, 1910, 1920,1930, 1940, 1950, 1960,1970,1980,1990,2000,2010,2020), labels = c( "1900", "1910","1920","1930","1940","1950","1960","1970","1980","1990","2000","2010","2020")) + scale_y_continuous(name ='Average years between Original and Remake') + ggtitle("Spacing between Movies and their Remakes") 
avg_gap_remakes
```
In both graphes we clearly see an upward trend, indicating the average length of time (for each year remakes and sequels were made) between a movie being produced and its sequel/remake has increased. This supports our prior assumptions, due to the fact that more time has occured since the first movies so there are more movies from longer ago to make a sequel or remake of. 

Our graphes show that on average, across all the years, less time occurs between an original and its sequel than an original and its remake. This could be due to the impatiance of many movie-viewers desiring a sequel soon after the original. But remakes are not sequentially linked to the orginal movie so time in between the original and remake could be longer. An alternative hypothesis for the average longer time between an original and remake could be due to the viewers not wanting to see a remake of a movie that was just produced. Perhaps, it is customary to give an original movie a large amount time before remaking it.

It is important to note, that due to IMDb not having informamtion in their "link_type" datatable specific to reboots, there are only sequels and remakes are shown in these graphes. There is the possiblity that IMDb's term "followed by", which I used as the sequel linking term, also includes reboots, so the movie reboots data may have been included in the sequels graph.  

In addition we see large spikes in both graphes around 2020, so be on the watch for upcoming remakes and sequels made from older movies. 


### Find information about the number of reboots, remakes, and sequels for each year

```{r}

remakes_reboots_sequels <- db %>%
  dbGetQuery(
    "SELECT kw.id, kw.keyword, t.title, t.production_year 
     FROM imdb.keyword kw
     JOIN movie_keyword mkw 
     ON kw.id = mkw.keyword_id
     JOIN title t
     ON t.id = mkw.movie_id
     WHERE kw.id IN (82, 1134, 20757) AND t.kind_id = 1;"

# 82, 1134, and 20757 are the keyword id numbers that correspond to the keywords remake, reboot, and sequel.
    
#This query is optimized because we use the primary keys in the keyword and title tables to join. When we use "where" to select the IDs from title and keyword, SQL only has to look at 1 and 3 rows, respectively. On movie keyword, keyword id is indexed, so we only have to look at 32 rows when referencing that table! Therefore, this query is extremely fast.
  )

remakes_by_year <- remakes_reboots_sequels %>%
  filter(!is.na(production_year)) %>%
  group_by(production_year) %>%
  summarize(
    reboot = sum(keyword == "reboot"),
    remake = sum(keyword == "remake"),
    sequel = sum(keyword == "sequel"))

remakes_by_year_gathered <- remakes_by_year %>%
  gather(key = "type", value = "count", reboot:sequel)



```

### Function: most of each type made in any single year
```{r}
most_in_year <- function(movie_type) {
  
  remakes_reboots_sequels %>%
  filter(keyword == movie_type, !is.na(production_year)) %>%
  group_by(production_year, keyword) %>%
  summarize(N = n()) %>%
  arrange(desc(N)) %>%
  head(1)
  
}

# The above function finds the year with the most movies of a given type made (e.g. the year with the most sequels)

list <- c("reboot", "remake", "sequel")
top_year_by_type <- lapply(list, FUN = most_in_year) %>% bind_rows()

```


### Graph the numbers of remakes, reboots, and sequels over time

```{r, warning = FALSE}

remakes_by_year_gathered$type <- factor(remakes_by_year_gathered$type, levels = c("sequel", "remake", "reboot"))

plot <- remakes_by_year_gathered %>%
  ggplot(aes(x = production_year, y = count, fill = type)) +
  geom_area(alpha = 0.6, position = "dodge") +
  scale_x_continuous(
    breaks = c(1900, 1910, 1920, 1930, 1940, 1950, 1960, 1970, 1980, 1990, 2000, 2010, 2020),
    expand = c(0,0)) + 
    # Add more labels (for each decade) to the plot
  geom_vline(xintercept = 2017) +
  # Add a line showing the current year. All movies after this line are planned and not actually finished.
  scale_fill_brewer(
    palette = "Set1",
    name= "Movie Type",
    breaks = c("sequel","remake", "reboot"),
    labels = c("Sequel", "Remake", "Reboot")) +
  xlab(NULL) +
  ylab("Number of Movies Produced") +
  ggtitle ("Movie Sequels, Remakes, and Reboots Since 1895")+
  geom_curve(
    x = 1983, xend = 1994,
    y = 105, yend = 116,
    arrow = arrow(length = unit(0.3,"cm")),
    curvature = -0.5) +
  geom_curve(
    x = 1950, xend = 1942,
    y = 78, yend = 67,
    arrow = arrow(length = unit(0.3, "cm")),
    curvature = -0.5) +
  geom_curve(
    x = 2008, xend = 2017,
    y = 15, yend = 5,
    arrow = arrow(length = unit(0.3, "cm")),
    curvature = 0.5) +
  geom_vline(xintercept = 2002, color = "red", linetype = 2) +
  geom_text(
    x = 2008, y = 100,
    label = "Spiderman\n released;\nleads to\nsuperhero\nmovie boom",
    color = "red", 
    size = 4) +
# For these vertical lines, we're matching color of the line to both the type of movie it affects most and to the text describing the event
  geom_vline(xintercept = 1927, color = "blue", linetype = 2) +
  geom_text(
    x = 1938, y = 110,
    label = "Movies add sound;\nmany silent films\nstart being remade",
    color = "blue",
    size = 4) +
  geom_vline(xintercept = 1948, linetype = 2) + 
  geom_text(
    x = 1958, y = 108,
    label = "US v. Paramount\n Pictures Case\nBreaks Up\nStudio System",
    size = 4) + 
# A black vertical line here means it pretty much equally affects all movie types
  geom_text(
    x = 1955, y = 80,
    label = "Most Remakes:\n67 in 1941",
    size = 4) +
  geom_text(
    x = 1980, y = 100,
    label = "Most Sequels:\n116 in 1994",
    size = 4) +
  geom_text(
    x = 2010, y = 20,
    label = "Most Reboots:\n5 in 2017",
    size = 4) 

plot +
  geom_point(data = top_year_by_type, aes(x = production_year, y = N), shape = 21, size = 1, fill = "white", color = "black") # Plot top years for each type of movie as dots on the plot

```
Above is a graph of the numbers of remakes, reboots, and sequels from 1895 to 2025. A solid vertical line is placed on the year 2017. All movies after that line have not been completed or released, and are just scheduled to be. Therefore, it is possible that those numbers will not be entirely accurate if film release dates are changed, or if any films are cancelled for some reason.

As we can see from the graph above, there *has* been an overall increase in all three types of movies over time, though only reboots are at their highest point in history right now. Both sequels and remakes have had higher peaks than they do now at earlier points in history (sequels peaked in 1994, and remakes in 1941). 

Certain historical events have affected the number of remakes, reboots, and sequels made, and several important ones are shown on the graph. The first large spike in movie remakes happened after 1927, when the first movies were made with sound. The addition of sound inspired filmmakers to remake many silent films from the past.

Then, there was a decrease in all types of films after 1948, when the supreme court case *United States v. Paramount Pictures, Inc.* resulted in the breakup of the studio system, as the court decided that the studios were in violation of US antitrust laws. When these studios were shut down, fewer movies overall were made, until the movie industry was able to regroup again and start producing more films, as seen by the later increase in sequels starting in the 1960s. 

The most recent large spike in sequels appears to be related to the rise in superhero movies. After Spiderman was released in 2002, there has been a massive boom of superhero franchises. Marvel and D.C. have put out huge numbers of movies featuring Batman, Ironman, Superman, Captain America, The Avengers, etc. Many of these are sequels in massive franchises, and as the public has grown to love these sequels, more and more are being made, so we are once again at a relatively high point in sequel production.

Reboots appear to be made on a much smaller scale than either sequels or remakes, with movies classified as "reboot" on IMDb only appearing within the last few years. Due to this recent increase in reboots, it is certainly possible that there will be a rise in that type of movie in the coming years.



The GitHub repository we used is linked [here](https://github.com/ambertin/mp3.git)

